---
title: "Fig S2"
author: "Cassandra"
date: "8/21/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
## for graphing
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())

## for genotype manipulation
library(glue)
library(ggtext)
```

# B, R, R+B FRRC

Examine rhythmicity of genotypes in all tested fluence rates of blue, red, and red + blue light. 

## load and format data 

### initial formatting

```{r}
# load data
rhythmicity <- read_csv("Dull_FRRC_rhythmicity_quint.csv")

# change factors 
rhythmicity <- rhythmicity %>% mutate(
  Experiment = as_factor(Experiment),
  Genotype = as_factor(Genotype),
  Condition = as_factor(Condition),
  RAE_class = as_factor(case_when(RAE <= 0.6 ~ "RAE < 0.6",
                        RAE > 0.6 ~ "RAE > 0.6",
                        is.na(RAE) ~ "none")))

# order RAE_class
rhythmicity$RAE_class = as_factor(rhythmicity$RAE_class) %>% 
  ordered(levels = c("none", "RAE > 0.6", "RAE < 0.6"))

summary(rhythmicity)
```

### summarize RAE class counts 

Calculate the number of plants for each RAE class when grouped by Genotype, Condition, Intensity, and Experiment. Also calculate the total number of plants across the three RAE groups. 

```{r}
# group to get counts for each RAE class
br_counts <- rhythmicity %>% group_by(Genotype, Condition, Intensity, RAE_class, Experiment) %>% 
  summarise(class_total = n())

# group to get total counts for each bar 
br_total <- rhythmicity %>% group_by(Genotype, Condition, Intensity, Experiment) %>% 
  summarise(n = n())

# join, calculate percentage
br_summary <- full_join(br_counts, br_total) %>% 
  mutate(percentage = class_total/n*100) 

summary(br_summary)
head(br_summary)

# average reps, calculate SD and SEM
br_reps <- br_summary %>% group_by(Genotype, Condition, Intensity, RAE_class) %>% 
  summarise(count = n(),
            avg_percentage = mean(percentage),
            sd = sd(percentage),
            sem = sd / sqrt(count))

summary(br_reps)

# include all classes so that bars are equal sizes when graphed
br_reps <- br_reps %>% complete(RAE_class)
summary(br_reps)
```

```{r}
# order RAE_class
br_reps$RAE_class = as_factor(br_reps$RAE_class) %>% 
  ordered(levels = c("RAE < 0.6", "RAE > 0.6", "none"))
```

### italicize genotypes 

**Set up italicized genotype labels.**   

```{r}
# italicize genotypes
br_reps_styled_tilt <- br_reps %>% mutate(
  gt = as_factor(ifelse(Genotype == "Col", "Col",
                        ifelse(Genotype == "cca1 lhy-100 rve468", 
                               "<i>cca1 lhy rve468-12</i>",
                               ifelse(Genotype == "cca1 lhy-100",
                                      "<i>cca1 lhy</i>",
                                      glue("<i>{Genotype}-11</i>"))))))

summary(br_reps_styled_tilt)

# order styled genotypes 
br_reps_styled_tilt$gt = as_factor(br_reps_styled_tilt$gt) %>% 
  ordered(levels = c("Col", "<i>rve468-11</i>", 
                     "<i>cca1 lhy</i>", 
                     "<i>cca1 lhy rve468-12</i>"))
summary(br_reps_styled_tilt$gt)
```

### prepare facet labels

```{r}
# set up light as more clear labels of Condition
br_reps_styled_tilt <- br_reps_styled_tilt %>% 
  mutate(light = as_factor(case_when(
    Condition == "cB" ~ "Blue",
    Condition == "cR" ~ "Red",
    Condition == "cR+cB" ~ "Red + Blue")))
summary(br_reps_styled_tilt$light)
```

## check data 

```{r}
# check out rhythmicity of cca1 lhy rve468 
br_reps_styled_tilt %>% 
  filter(Genotype == "cca1 lhy-100 rve468") %>% 
  filter(RAE_class == "RAE < 0.6")
```

### check seedling count for each rep

```{r}
# for supplementary figure 
rhythmicity %>% 
  group_by(Condition, Intensity, Experiment, Genotype) %>% 
  tally() %>% 
  print(n = 135)

rhythmicity %>% 
  group_by(Condition, Intensity, Experiment, Genotype) %>% 
  tally() %>% 
  summary()
```

All reps included. There are 3 reps for everything in cB. There are 3 reps for everything in cR except cca1 lhy at 1 uE and only 2 reps at 200 uE. There are 3 reps for everything in cR+cB except quint at 1 uE and 10 uE, and rve468 and cca1 lhy at 200 uE. 

## create figure

### bar graph with error bars

```{r}
# figure specifications
facet_label_size <- 18
axis_text_size = 16
axis_title_size = 18
legend_text_size <- 16
legend_title_size <- 18
legend_symbol_size <- unit(2, "line")
point_size = 1.75
point_alpha = 0.5
point_stroke = 1
errorbar_width <- 0.25
errorbar_size <- 0.75
```

```{r fig.width = 14, fig.height = 8}
# save to pdf 
#pdf("supp2_bar_sem_tilt.pdf", width = 14, height = 8, bg = "white")

br_reps_styled_tilt %>%  
  ggplot(aes(gt, avg_percentage, fill = RAE_class)) +
  geom_bar(position = position_dodge(), stat = "identity") +
  facet_grid(light ~ Intensity) +
  geom_errorbar(aes(ymin = avg_percentage - sem, 
                    ymax = avg_percentage + sem),
                width = errorbar_width, linewidth = errorbar_size,
                position = position_dodge(0.9)) +
  labs(y = "Percentage of Plants") + 
  scale_fill_manual(values = c("none" = "firebrick2", 
                               "RAE > 0.6" = "gold", 
                               "RAE < 0.6" = "mediumblue"), name = "RAE") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_markdown(size = axis_text_size, angle = 45, hjust = 1),
        axis.title.y = element_text(size = axis_title_size),
        axis.text.y = element_text(size = axis_text_size),
        strip.text = element_text(size = facet_label_size),
        legend.text = element_text(size = legend_text_size),
        legend.title = element_text(size = legend_title_size),
        legend.position = "top",
        legend.justification = "center",
        legend.key.size = legend_symbol_size)

#dev.off()
```
