---
title: "Fig S3"
author: "Cassandra"
date: "8/21/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
## for modeling
library(lme4)
library(lmerTest)
## for qqmath to get QQ plot
library(lattice)
## for tidy to get p-values
library(broom)
## for Tukey testing and visualization 
library(emmeans)
library(multcomp)
library(multcompView)

## for graphing
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
## for figure arrangement
library(gridExtra)

## for genotype manipulation
library(glue)
library(ggtext)
```

# B, R, R+B, DD Period Boxplots

Compare period between genotypes in constant darkness, constant 10 uE red, constant 10 uE blue, and constant 10 uE red + blue light.

## load data 

```{r}
# load B, R, R+B data
rhythmicity <- read_csv("Dull_FRRC_rhythmicity_quint.csv")

# assign reps by experiment
rhythmicity <- rhythmicity %>% mutate(
  rep = as_factor(case_when(
    Experiment == "Dull617" | 
      Experiment == "Dull619" |
      Experiment == "Dull620" |
      Experiment == "Dull621" |
      Experiment == "Dull625" ~ "1",
    Experiment == "Dull622" |
      Experiment == "Dull624" |
      Experiment == "Dull626" ~ "2",
    Experiment == "Dull632" | 
      Experiment == "Dull633" |
      Experiment == "Dull634" ~ "3")))

summary(rhythmicity$rep)
```

```{r}
# load dark data 
dark <- read_csv("Dull_DD_rhythmicity_quint.csv")

# add condition and intensity columns
dark <- dark %>% mutate(
  Condition = as_factor("DD"),
  Intensity = 0
)

# assign reps by experiment
dark <- dark %>% mutate(
  rep = as_factor(case_when(
    Experiment == "Dull603" ~ "1",
    Experiment == "Dull606" ~ "2",
    Experiment == "Dull607" ~ "3")))
  
summary(dark$rep)
```

## format data 

```{r}
# join together
all <- full_join(rhythmicity, dark)

# change factors 
all <- all %>% mutate(
  Experiment = as_factor(Experiment),
  Genotype = as_factor(Genotype),
  Condition = as_factor(Condition),
  RAE_class = as_factor(case_when(RAE <= 0.6 ~ "RAE < 0.6",
                        RAE > 0.6 ~ "RAE > 0.6",
                        is.na(RAE) ~ "none")))

summary(all)
```

```{r}
# order RAE_class
all$RAE_class = as_factor(all$RAE_class) %>% 
  ordered(levels = c("none", "RAE > 0.6", "RAE < 0.6"))

summary(all)
```

```{r}
# filter for only rhythmic plants
supp3 <- all %>%
  filter(Intensity == 10 | Condition %in% c("DD")) %>% 
  filter(RAE <= 0.6) %>% 
  droplevels()

# change light values for better labels
supp3 <- supp3 %>% 
  mutate(light = as_factor(case_when(
    Condition == "cB" ~ "Blue",
    Condition == "cR" ~ "Red",
    Condition == "cR+cB" ~ "Red + Blue",
    Condition == "DD" ~ "Dark")))

# order light values to match fig 1 layout
supp3$light = as_factor(supp3$light) %>% 
  ordered(levels = c("Dark", "Red",
                     "Red + Blue", "Blue"))

summary(supp3)
```

## italicize genotypes 

**Set up italicized genotype labels.**   

```{r}
supp3_styled <- supp3 %>% mutate(
  gt = as_factor(ifelse(Genotype == "Col", "Col",
                        ifelse(Genotype == "cca1 lhy-100 rve468", 
                               "<i>cca1<br>lhy<br>rve468-12</i>",
                               ifelse(Genotype == "cca1 lhy-100",
                                      "<i>cca1<br>lhy</i>",
                                      glue("<i>{Genotype}-11</i>"))))))

# order styled genotypes 
supp3_styled$gt = as_factor(supp3_styled$gt) %>% 
  ordered(levels = c("Col", "<i>rve468-11</i>", 
                     "<i>cca1<br>lhy</i>", 
                     "<i>cca1<br>lhy<br>rve468-12</i>"))
summary(supp3_styled$gt)
```

## set aesthetic scales

Set shapes by genotype. 

```{r}
quint_shapes <- c("Col" = 16, "<i>rve468-11</i>" = 1,
                        "<i>cca1<br>lhy</i>" = 17, 
                        "<i>cca1<br>lhy<br>rve468-12</i>" = 2)
```

Set colors by genotype with allele differentiation. 

```{r}
quint_style_colors <- c("Col" = "#1F78B4", "<i>rve468-11</i>" = "#1B9E77",
                        "<i>cca1<br>lhy</i>" = "#E31A1C", 
                        "<i>cca1<br>lhy<br>rve468-12</i>" = "#6A3D9A")
```

## initial boxplot 

```{r fig.height = 5, fig.width = 7}
# all reps together
supp3_styled %>% 
  filter(RAE <= 0.6) %>%
  ggplot(aes(gt, Period)) + 
  geom_boxplot(outlier.size = -1) + 
  geom_point(aes(color = gt), alpha = 0.8, position = position_jitterdodge()) +
  facet_wrap(~ light) + 
  scale_color_manual(values = quint_style_colors) + 
  theme(axis.text.x = element_markdown(),
        axis.title.x = element_blank(),
        legend.position = "none")
```

## stats with lmerTest

Want to examine period ~ genotype. Include experiment or rep as a random effect.  

### check models

```{r fig.show = "hold", out.width = "50%"}
# pull out all conditions that need subsetting (light works for this)
conditions <- unique(supp3_styled$light)
conditions

# run loop over all quality and intensity subsets
for (i in 1:length(conditions)){
  # subset df
  df <- supp3_styled[supp3_styled$light == conditions[i],]
  # create model 
  model <- lmer(Period ~ Genotype + (1|rep), data = df)
  # set the figures side by side 
  par(mfrow=c(1,2))
  # plot normality of residuals
  print(qqmath(model, main = as.character(conditions[i])))
  # plot equality of variances
  print(plot(model,
       sqrt(abs(resid(.)))~fitted(.),
     type=c("p","smooth"), col.line=1,
     main = as.character(conditions[i])))
}
```

**Normality is okay but not great. Red + Blue variance looks okay.**

### ANOVA

```{r}
# create results data frame
aov_res <- data.frame(matrix(NA, 
                             nrow = length(1:length(conditions)),
                             ncol = 8))
# assign column names to match tidied Tukey output
colnames(aov_res) <- c("light", "term", "sumsq", "meansq", 
                       "NumDF", "DenDF", "statistic", "p.value")

# run loop over all light subsets
for (i in 1:length(conditions)){
  # subset df
  df <- supp3_styled[supp3_styled$light == conditions[i],]
  # create model 
  model <- lmer(Period ~ Genotype + (1|rep), data = df)
  # specify condition value in results
  aov_res$light[i] <- as.character(conditions[i])
  # run anova
  aov <- anova(model)
  # tidy anova into tibble
  tidied <- tidy(aov)
  # assign tidied variables into results data frame
  aov_res$term[i] <- tidied$term
  aov_res$sumsq[i] <- tidied$sumsq
  aov_res$meansq[i] <- tidied$meansq
  aov_res$NumDF[i] <- tidied$NumDF
  aov_res$DenDF[i] <- tidied$DenDF
  aov_res$statistic[i] <- tidied$statistic
  aov_res$p.value[i] <- tidied$p.value
}

aov_res %>% 
  select(!c("sumsq":"DenDF")) %>% 
  arrange(light)
```

**Significant difference in genotype for all of these (Dark, 10B, 10R, 10R+B).**

### Tukey

```{r}
# run loop over all light subsets
for (i in 1:length(conditions)){
  # subset df
  df <- supp3_styled[supp3_styled$light == conditions[i],]
  # create model 
  model <- lmer(Period ~ Genotype + (1|rep), data = df)
  # run emmeans 
  emm <- emmeans(model, ~ Genotype)
  # add line of space between results + label condition of tested subset
  cat(paste("", 
            as.character(conditions[i]),
            "", sep = "\n"))
  # print Tukey results in table
  print(pwpm(emm, diffs = FALSE))
}
```

### convert Tukey results into letter groups

Want to know which groups are significantly different from each other, and which groups are not significantly different. 

```{r}
# create results data frame
tukey_res <- data.frame(matrix(NA, 
                               nrow = length(1:(length(conditions)*4)),
                               ncol = 8))
# assign column names to match output
colnames(tukey_res) <- c("light", "Genotype", "emmean", "SE", 
                         "df", "lower.CL", "upper.CL", ".group")

# run loop over all light subsets
for (i in 1:length(conditions)){
  # subset df
  df <- supp3_styled[supp3_styled$light == conditions[i],]
  # create model 
  model <- lmer(Period ~ Genotype + (1|rep), data = df)
  # run emmeans 
  emm <- emmeans(model, ~ Genotype)
  # assign groups  
  res <- cld(emm, level = 0.95, Letters = letters)
  for (j in 1:4){
    # set row number based on i and j
    x <- (4*(i-1)) + j
    # specify condition value in results 
    tukey_res$light[x] <- as.character(conditions[i])
    #assign variables into results data frame
    tukey_res$Genotype[x] <- as.character(res$Genotype[j])
    tukey_res$emmean[x] <- res$emmean[j]
    tukey_res$SE[x] <- res$SE[j]
    tukey_res$df[x] <- res$df[j]
    tukey_res$lower.CL[x] <- res$lower.CL[j]
    tukey_res$upper.CL[x] <- res$upper.CL[j]
    tukey_res$.group[x] <- res$.group[j]
  }
}

head(tukey_res)

# adjust letter format to plot
tukey_res <- tukey_res %>% 
  mutate(
    light = as_factor(light),
    Genotype = as_factor(Genotype),
    group = sapply(.group, str_remove_all, " "))

# view results
tukey_res %>% 
  select(!c(df:upper.CL))
```

## format data 

Calculate the maximum data point for each condition and genotype so that Tukey test lettering is displayed at the appropriate position.  

```{r}
# calculate max period for each gt and condition
max_per <- supp3_styled %>% group_by(light, Genotype) %>% 
  summarise(max_per = max(Period))

# order light in Tukey results for joining
tukey_res$light = as_factor(tukey_res$light) %>% 
  ordered(levels = c("Dark", "Red", "Red + Blue", "Blue"))

# join max period into the Tukey results 
(tukey_res <- full_join(tukey_res, max_per))
```

## italicize genotypes in Tukey data frame

```{r}
# set up italicized genotypes
tukey_res_styled <- tukey_res %>% mutate(
  gt = as_factor(ifelse(Genotype == "Col", "Col",
                       ifelse(Genotype == "cca1 lhy-100 rve468", "<i>cca1<br>lhy<br>rve468-12</i>",
                       ifelse(Genotype == "cca1 lhy-100", "<i>cca1<br>lhy</i>",
                    glue("<i>{Genotype}-11</i>"))))))

# order styled genotypes 
tukey_res_styled$gt = as_factor(tukey_res_styled$gt) %>% 
  ordered(levels = c("Col", "<i>rve468-11</i>", 
                     "<i>cca1<br>lhy</i>", 
                     "<i>cca1<br>lhy<br>rve468-12</i>"))
summary(tukey_res_styled)
```


## final boxplot 

```{r}
# size specifications
point_size = 2
point_alpha = 0.8
axis_text_size = 14
axis_title_size = 16
facet_label_size = 16
cld_size = 5
```

```{r fig.height = 5, fig.width = 9}
# save supp fig3 to pdf
#pdf("supp3.pdf", width = 9, height = 5, bg = "white")

# all reps together
supp3_styled %>% 
  ggplot(aes(gt, Period)) + 
  geom_boxplot(outlier.size = -1) + 
  geom_point(aes(color = gt, shape = gt), 
             alpha = point_alpha, size = point_size,
             position = position_jitterdodge()) +
  geom_text(data = tukey_res_styled, aes(x = gt, y = (max_per + 3.5),
                                         label = group),
            color = "black", size = cld_size, show.legend = FALSE) + 
  facet_wrap(~ light) + 
  scale_y_continuous(limits = c(13, 39), breaks = c(seq(15, 35, 5))) + 
  background_grid(major = 'y', minor = "none")  +
  scale_colour_manual(values = quint_style_colors) + 
  scale_shape_manual(values = quint_shapes) + 
  theme(axis.text.x = element_markdown(size = axis_text_size),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = axis_text_size),
        axis.title.y = element_text(size = axis_title_size),
        strip.text = element_text(size = facet_label_size),
        legend.position = "none")

#dev.off()
```
